Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-01 paper/manuscript.tex
Start working on last results section

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/sequence_preference/sequence_preference.pdf}}
\caption{\label{fig:sequence_preference}
{\bf The HA sequences of trunk nodes tend to be more preferred than those of side branch nodes.}
We used the preferences to calculate the HA sequence preference of every node in a human H3N2 phylogenetic tree.
To calculate the sequence preference, we used the entire HA sequence (top panel), only epitope sites (center panel) or non-epitope sites (bottom panel) as defined by~\cite{wolf2006long}.
Higher preferences are closer to zero while lower preferences are more negative.
The preferences at all sites and at epitope sites increase as the tree approaches the Perth/2009 strain, indicating epistasis among epitope sites.
The trunk nodes generally exhibit higher sequence preferences than do side branch nodes.
}
\end{figure}

@@

\subsection*{The H1 and H3 preferences at many sites have shifted}
How shifted are the preferences between evolutionarily distant homologs such as H1 and H3?
Although we have previously compared the preferences between related protein homologs of NP~\citep{doud2015site} and of Env \comment{cite haddox2017}, the HA homologs we have experimentally studied are considerably more diverged than either of these pairs.
The WSN/1933 H1 and the Perth/2009 H3 fall into two phylogenetically distinct groups and share 42\% amino-acid identity (Figure~\ref{fig:distance_distribution}A).
\begin{itemize}
\item Correlations are higher within homolog replicate than between homologs
\item Next aligned and compared preferences using a correction method to account for experimental noise
\end{itemize}

>>>>

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/sequence_preference/sequence_preference.pdf}}
\caption{\label{fig:sequence_preference}
{\bf The HA sequences of trunk nodes tend to be more preferred than those of side branch nodes.}
We used the preferences to calculate the HA sequence preference normalized by the number of sites of every node in a human H3N2 phylogenetic tree.
To calculate the sequence preference, we used the entire HA sequence (top panel), only epitope sites (center panel) or non-epitope sites (bottom panel) as defined by~\cite{wolf2006long}.
Higher preferences are closer to zero while lower preferences are more negative.
The preferences at all sites and at epitope sites increase as the tree approaches the Perth/2009 strain, indicating epistasis among epitope sites.
The trunk nodes generally exhibit higher sequence preferences than do side branch nodes.
}
\end{figure}

@@

\subsection*{The H1 and H3 preferences at many sites have shifted}
How shifted are the preferences between evolutionarily distant homologs such as H1 and H3?
Although we have previously compared the preferences between related protein homologs of NP~\citep{doud2015site} and of Env \comment{cite haddox2017}, the HA homologs we have experimentally studied are considerably more diverged than either of these pairs.
The WSN/1933 H1 and the Perth/2009 H3 fall into two phylogenetically distinct groups and share 42\% amino-acid identity (Figure~\ref{fig:distance_distribution}A), providing an opportunity to analyze what has shifted and what has remained conserved.
Simply correlating the preferences between H1 and H3 reveals that the replicate measurements within a homolog are more correlated than between homologs (Figure~\ref{fig:distance_distribution}B).

However, to quantify the extent of mutational shifts at a site-by-site level, we used an approach described in \comment{haddox2017}.
We aligned Perth/2009 H3 and WSN/1933 H1 and at each alignable site calculated the difference in preferences between homologs while correcting for experimental noise within homologs.
The distribution of shifts is shown in Figure~\ref{fig:distance_distribution}C.
